---
title: "hw04"
author: "Xinzhe Dong"
date: "October 4, 2017"
output: 
 github_document 
---

### load the packages

```{r echo=FALSE}
library(gapminder)
library(tidyverse)
library(knitr)
library(kableExtra)
library(ggplot2)
library(forcats)
library(pander)
```


### 1. General data reshaping and relationship to aggregation




### 2. Join, merge, look up
Create a second data frame, complementary to Gapminder. Join this with Gapminder using a  dplyr join function and make some observations about the process and result. Explore the different types of joins. 

**The second data frame**:

One row per continent("Asia", "Oceania", "Europe", "Antarctica"), a continent variable and two more variable with extra info. One is areas of continents, in square miles, and another one is northern versus southern hemisphere.

```{r}
(extra_infor <- data.frame(continent=c("Asia", "Oceania", "Europe", "Antarctica"),
                     areas=c(17139445, 2967909, 3997929, 5405000),
                     hemisphere = c("northern", "southern", "northern", "southern")) %>% 
     as.tbl)
```

**1. To the `gapminder` dataset, add the "areas" and "hemisphere" variable using the `extra_infor` tibble.**

```{r}
gapminder%>%
  left_join(extra_infor, by = "continent")
```


**2. To the `gapminder` dataset, add the "areas" and "hemisphere" variable using the `extra_infor` tibble, but only keeping obervations for which areas and hemisphere are available. **


```{r}
gapminder%>%
  inner_join(extra_infor, by = "continent")
```

**3. Output the rows in `extra_infor` corresponding to continents that _are not_ found in the `gapminder` dataset. **


```{r}
extra_infor%>%
  anti_join(gapminder, by = "continent")
```

**4. Output the rows in `extra_infor` corresponding to continents that _are_ found in the `gapminder` dataset. **


```{r}
extra_infor%>%
  semi_join(gapminder, by = "continent")
```


**5. Construct a tibble that joins `gapminder` and `extra_infor`, so that all rows found in each original tibble are also found in the final tibble. **


```{r}
full_join(gapminder,extra_infor, by = "continent")
```


**6. Subset the `gapminder` dataset to have continents that are only found in the `extra_infor` data frame. **

```{r}
gapminder%>%
  semi_join(extra_infor)
```


**7. Subset the `gapminder` dataset to have continents that are _not_ found in the `extra_infor` data frame. **

```{r}
gapminder%>%
  anti_join(extra_infor)
```