---
title: "hw 02"
author: "Xinzhe Dong"
date: "September 22, 2017"
output: 
  html_document: 
    keep_md: yes
---

```{r}
library(gapminder)
library(tidyverse)
library(knitr)
library(kableExtra)

gapminder
str(gapminder)
```
1. It is a data.frame.
2. Classes ‘tbl_df’, ‘tbl’ and 'data.frame'
3. 6 variables/columns
4. 1704 rows/obs.
5. I can use str(), ncol(), nrow(), length(), dim(), names().
6. Country is factor, continent is factor, year is integer, lifeExp is number, pop is integer, gdpPercap is num.

```{r}
cont <- gapminder$continent
class(cont)
summary(cont)
levels(cont)
nlevels(cont)

lifeex <- gapminder$lifeExp
summary(lifeex)
```

Let’s try to make a basic scatterplot of year vs. lifeExp.
```{r}
ggplot(gapminder, aes(x=year, y=lifeExp)) +
    geom_point()
```

histogram of lifeExp
```{r}
ggplot(gapminder, aes(x = lifeExp)) +
  geom_histogram()
```

A boxplot of lifeExp and one continents
```{r}
ggplot(gapminder, aes(x = continent, y = lifeExp)) + geom_boxplot()

```

Use filter() to create data subsets that you want to plot.
```{r}
filter(gapminder, year <= 1999, year >= 1960)

```

Practice piping together filter() and select(). Possibly even piping into ggplot().
```{r}
gapminder%>%
 select(year, lifeExp)%>%
filter(year <= 1999, year >= 1960)%>%
  ggplot(aes(x=year, y=lifeExp)) +
    geom_point()

```

This only gives part of the data.
```{r}
filter(gapminder, country == c("Rwanda", "Afghanistan"))

```

```{r}
filter(gapminder, (country=="Rwanda"|country=="Afghanistan"))

```

```{r}
filter(gapminder, country %in% c("Rwanda", "Afghanistan"))

```

```{r}
gapminder%>%
    filter(country == "Australia")%>%
  knitr::kable()
```
