---
title: "hw 06"
author: "Xinzhe Dong"
date: "November 2, 2017"
output: github_document
---

### Abstract

We are going to 


#### Table of content
-   [Load data and packages](#load-data-and-packages)
-   [Character data](#character-data)
-   [Writing functions](#writing-functions)


Load data and packages
----------------------

[*Back to the top*](#abstract)

```{r message=FALSE}
library(tidyverse)
library(stringr)
library(glue)
library(gapminder)
library(ggplot2)
library(MASS)

```

Character data
--------------

[*Back to the top*](#abstract)

**Exercise 1** Use `str_length()` and `str_sub()` to extract the middle character from a string. If the string has an even number of characters, extract the middle two characters.
```{r}
my_string1 <- "statistics"  # a string with an even number of characters
my_string2 <- "hello"       # a string with an odd number of characters 

# a function string_extract() which takes in a string and return the middle character(s) from a string
string_extract <- function(x){
  x_length <- str_length(x)
  if((x_length %% 2) == 0) {
  str_sub(x, start=x_length/2, end=x_length/2+1)} 
  else {
  str_sub(x, x_length/2+0.5, end=x_length/2+0.5)} 
}

# let's test our function
string_extract(my_string1)
string_extract(my_string2)
```


**Exercise 2** What does `str_trim()` do? What’s the opposite of str_trim()?

```{r}
# str_trim()
str_trim("  hello November  ")
str_trim("  hello November  ", side = "left")

# str_pad()
str_pad("hello November",width=20,side = "left")
str_pad("hello November",width=30,side = "both")

```

```{r}

```

**We can tell that:** 

* `str_trim()` trim whitespace from start and end of string.

* The opposite of `str_trim()` is `str_pad()`, it is used to add whitespace. 



Writing functions
-----------------

[*Back to the top*](#abstract)

**Summary:** I will write two functions in this section, `le_lin_fit` and `le_r_fit`. The input and output for both functions are summarised as follow:

**Input:** both function take in a data.frame that contains (at least) a lifeExp variable and a variable for year

**Output:** 

* `le_lin_fit`: a list of estimated intercept and slope, with the estimated mspe, from a linear regression of lifeExp on year

* `le_r_fit`: a list of estimated intercept and slope, with the estimated mspe, from a robust regression of lifeExp on year

**(1)** Get a subset of gapminder so that it only contains entries of China. And plot the lifeExp against year.
```{r}
j_country <- "China"
j_dat <- gapminder %>% 
  filter(country == j_country)

p <- ggplot(j_dat, aes(x = year, y = lifeExp))
p + geom_point() + geom_smooth(method = "lm", se = FALSE)
```

**(2)** Split data into a training and a test set.
```{r}
set.seed(123456)
n <- nrow(j_dat)
ii <- sample(n, floor(n/4))
dat.te <- j_dat[ ii, ]
dat.tr <- j_dat[ -ii, ]
```

**(3)** Fit a linear model to the training set. And calculate the estimated mspe on the test set. 
```{r}
le_lin_fit <- function(dat, offset = 1952) {
  the_fit <- lm(lifeExp ~ I(year - offset), dat)
  # predict on the test set
  pr.to <- predict(the_fit, newdata=dat.te)
  mspe <- with(dat.te, mean((dat.te$lifeExp - pr.to)^2) )
  return(list(the_fit,"mspe"=mspe))
}
le_lin_fit(dat.tr)
```

**(4)** Fit a robust regression model to the training set. And calculate the estimated mspe on the test set. 
```{r}
le_r_fit <- function(dat, offset = 1952) {
  the_fit <- MASS::rlm(formula = lifeExp ~ I(year - offset), data=dat)
  # predict on the test set
  pr.to <- predict(the_fit, newdata=dat.te)
  mspe <- with(dat.te, mean((dat.te$lifeExp - pr.to)^2) )
  return(list(the_fit,"mspe"=mspe))
}

le_r_fit(dat.tr)
```


**Note:**

* The two regression functions return slightly different intercept and slopes.

* The robust regression model returns a smaller estimated mspe, which means it is more robust to the existence of outliers.

* The training set has 9 observations while the test set has 3 observations. Therefore the sample size is relatively small to make some concrete conclusions. However, the major point of this task is to recognize that the robust regression model is less likely to be influenced by extreme values.

