---
title: "hw05"
author: "Xinzhe Dong"
date: "October 12, 2017"
output: github_document
---

### Abstract

We are going to explore the gapminder data set. We will be practicing data reshaping and join.

#### Table of content
-   [Load data and packages](#load-data-and-packages)
-   [General data reshaping and relationship to aggregation](#general-data-reshaping-and-relationship-to-aggregation)
-   [Join, merge, look up](#join-merge-look-up)

Load data and packages
----------------------

[*Back to the top*](#abstract)

```{r}
library(gapminder)
library(tidyverse)
library(knitr)
library(kableExtra)
library(ggplot2)
library(forcats)
library(pander)
```


Factor management

Get to know your factor before you start touching it!
```{r}
fct_count(gapminder$continent)
```

Drop all the rows corresponding to "Oceania".
```{r}
h_gap <- gapminder %>%
  filter(continent != "Oceania")

fct_count(h_gap$continent)
nlevels(h_gap$continent)
```
We dropped 24 rows associated with "Oceania". However, the levels of the continent factor itself do not change. "Oceania" is still a level of "continent".

Dropping "Oceania" level.
```{r}
h_gap_dropped <- h_gap %>% 
  droplevels()
fct_count(h_gap_dropped$continent)
nlevels(h_gap_dropped$continent)
```
Now, we dropped 24 rows associated with "Oceania". Additionally, we also dropped the level "Oceania" of the factor "continent".


Change order of the levels

Reorder Gapminder continents by lifeExp
```{r}
## order continent according to minimum life exp
fct_reorder(h_gap_dropped$continent, h_gap_dropped$lifeExp, min) %>% 
  levels() %>% head()

h_gap_dropped$continent<-fct_reorder(h_gap_dropped$continent, h_gap_dropped$lifeExp, min)

h_gap_dropped%>%
  group_by(continent)%>%
  summarise(min(lifeExp))%>%
  kable()
```

Common part:
**While you’re here, practice writing to file and reading back in (see next section).**
```{r}
write_csv(h_gap_dropped, "h_gap_dropped.csv")
read_csv("h_gap_dropped.csv")
```


**Characterize the (derived) data before and after your factor re-leveling.**

Explore the effects of arrange().

```{r}
gap_asia_2007 <- gapminder %>% filter(year == 2007, continent == "Asia")
# unarranged, reordered
ggplot(gap_asia_2007, aes(x = lifeExp, y = country)) + geom_point() +
        labs(title= "Plot of lifeExp of countries in Asian in 2007") 

# arranged
gap_asia_2007%>%
  arrange(country, lifeExp)%>%
    ggplot(aes(x = lifeExp, y = country)) + geom_point() +
        labs(title= "Plot of lifeExp of countries(arranged) in Asian in 2007") 

# reordered
ggplot(gap_asia_2007, aes(x = lifeExp, y = fct_reorder(country, lifeExp, min))) +
  geom_point() + labs(title= "Plot of lifeExp of countries(reordered) in Asian in 2007") 


# first reordered, then arranged
gap_asia_2007%>%
  mutate(country=fct_reorder(country, lifeExp, min))%>%
  arrange(country, lifeExp)%>%
    ggplot(aes(x = lifeExp, y = country)) + geom_point() +
        labs(title= "Plot of lifeExp of countries(reordered, arranged) in Asian in 2007") 
```

**Note**: 

* Merely arranging the data has no effect on a figure.
* Reordered factor levels makes plots much better.


Explore the effects of reordering a factor and factor reordering coupled with arrange(). Especially, what effect does this have on a figure?
These explorations should involve the data, the factor levels, and some figures.

