---
title: "hw03"
author: "Xinzhe Dong"
date: "September 26, 2017"
output: 
  github_document
---

### load the packages

```{r echo=FALSE}
library(gapminder)
library(tidyverse)
library(knitr)
library(kableExtra)
library(ggplot2)
library(forcats)
library(pander)
```


### 1. Get the maximum and minimum of GDP per capita for all continents.

```{r}
gapminder %>% 
  group_by(continent) %>% 
  summarize(min_gdpPercap=min(gdpPercap), max_gdpPercap=max(gdpPercap), range = max_gdpPercap - min_gdpPercap)
```

#### Plot the max and min of GDP per capita for all continents.
```{r}
gapminder %>% 
  group_by(continent) %>% 
  summarize(min_gdpPercap=min(gdpPercap), max_gdpPercap=max(gdpPercap))%>%
  ggplot(aes(x = continent)) + 
  geom_point(aes(y = min_gdpPercap)) +
  geom_point(aes(y = max_gdpPercap)) +
  labs(y="Max & Min gdpPercap", title= "Plot of max & min of gdpPercap for all continents") 
```

#### We observe the followings:

* Both the maximum and minimum of GDP per capita for Africa is the smallest among the five continents.

* Asia has the largest range of GDP per capita, and the largest maximum of GDP per capita.

* Oceania has the largest minimum gdpPercap.



### 2. Look at the spread of GDP per capita within the continents.

```{r}
# Statistical summary of gdpPercap within continents
gapminder%>%
  group_by(continent)%>%
  summarize(min_gdpPercap=min(gdpPercap), first_q=quantile(gdpPercap)[2], median_gdpPercap=median(gdpPercap), third_q=quantile(gdpPercap)[4], max_gdpPercap=max(gdpPercap))

# A density plot of gdpPercap
ggplot(gapminder, aes(x = gdpPercap, color = continent)) + geom_density() +
    labs(title= "Density plot of gdpPercap within the continents") 


# boxplots of GDP per capita for all continents. 
ggplot(gapminder, aes(x = continent, y = gdpPercap)) + geom_boxplot() +
      labs(title= "Boxplot of gdpPercap within the continents") 

```

#### From the above two plots, we observe that:

* Oceania has the largest median of gdpPercap among the five continents. Europe has the second largest median of gdpPercap among the five continents.

* Africa has the smallest median of gdpPercap among the five continents.

* The gdpPercap of Asia is more widely spreaded than the other continents.

* The gdpPercap of the five countries are all skewed to the right. The distribution  of gdpPercap of Oceania is closer to the normal distribution. 



### 3. Compute a weighted mean, weighting by population, of life expectancy for different years. 

```{r}
gapminder %>% 
  group_by(year) %>% 
  summarize(mean_lifeex=weighted.mean(lifeExp,pop))
```

#### Plot the weighted mean of life expectancy vs. year.
```{r}
gapminder %>% 
  group_by(year) %>% 
  summarize(mean_lifeex=weighted.mean(lifeExp,pop)) %>%
  ggplot(aes(x = year, y = mean_lifeex)) + geom_line() +
      labs(title= "Plot of weighted mean of life expectancy vs. year") 

```

#### From the plot, we can tell that:

The weighted mean of life expectancy from 1952 to 2007 is increasing at an approximately constant rate. 



### 4. How is life expectancy changing over time on different continents?
#### Calculate the mean life expectancy over time on different continents
```{r}
gapminder %>%
  group_by(continent,year)%>%
  summarize(mean_lifeExp = mean(lifeExp))
```

#### Plot life expectancy vs. year on different continents
```{r}
p1 <- ggplot(data = gapminder) + 
  geom_point(mapping = aes(x = year, y = lifeExp)) + 
  facet_wrap(~ continent, nrow = 2) + 
        labs(title= "Plot of life expectancy over time on different continents") 

p1 + theme_bw() +
  theme(strip.background = element_rect(fill="orange"),
          axis.title = element_text(size=14),
          strip.text = element_text(size=14, face="bold"))
```

#### We can tell that:

The life expectancy is increasing over time on all the continents.



### 5. Report the absolute and/or relative abundance of countries with low life expectancy over time by continent: 

Compute a mean of life expectancy worldwide. Then determine how many countries on each continent have a life expectancy less than the mean life expectancy, for each year.

```{r}
average_age <- mean(gapminder$lifeExp)
gapminder %>% 
  filter(lifeExp<average_age) %>%
    group_by(continent,year) %>% 
    summarize(No._countries_with_low_lifeExp=length(country))
```

#### Note:

* If there is no value for a continent on that year. then it means there is no country that year on that continent has a life expectancy less than the mean life expectancy.

* For example: There is no value for the continent "Oceania" in any year, which means Oceania does not have any country which has a life expectancy less than the mean life expectancy during the whole data collection period.

#### Plot the number of countries with low life expectancy over time by continent.

```{r}
gapminder %>% 
  filter(lifeExp<average_age) %>%
    group_by(continent,year) %>% 
    summarize(No._countries_with_low_lifeExp=length(country)) %>%
    ggplot(aes(x = year, y = No._countries_with_low_lifeExp)) + geom_line() +
      facet_wrap(~ continent, nrow = 2) +
          labs(title= "Plot of number of countries with low life expectancy over time by continent") 

```

#### From the plots, we can tell that:

* The number of countries with low life expectancy is decreasing over time in each continent.

* There is not plot for Oceania, because Oceania does not have any country which has a life expectancy less than the mean life expectancy during the whole data collection period.

* For Americas and Europe, the number of countries with low life expectancy decreases to zero after a time, so the line on their graph is shorter compare to the line on the graph of Africa and Asia.


### 6. But I want to do more!


#### (1) Get table and figure side-by-side
<style type="text/css">
.twoC {width: 100%}
.clearer {clear: both}
.twoC .table {max-width: 50%; float: right}
.twoC img {max-width: 50%; float: left}
</style>

```{r}
library(ggplot2)
suppressMessages(library(dplyr))
#devtools::install_github("jennybc/gapminder")
library("gapminder")
gtbl <- tbl_df(gapminder)
```

<div class="twoC">
```{r results = 'asis', echo=FALSE}
tbl1 <- gtbl %>%
  group_by(continent) %>%
  summarise_each(funs(mean, median), gdpPercap)
knitr::kable(tbl1)
```

```{r echo=FALSE}
ggplot(gtbl, aes(x = continent, y = gdpPercap, color = continent)) +
  geom_point() + ggtitle("GDP per cap by continent") 
```
</div>
<div class="clearer"></div>

#### (2) Try a table helper package (pander)

```{r}
gapminder %>%
  filter(country=="Australia", year %in% c(1952,2007 ))%>%
  pandoc.table(style = "grid", caption = "Entries of Australia in 1952 and 2007")
```
