---
title: "hw10"
author: "Xinzhe Dong"
date: "11/30/2017"
output: github_document
---

### Abstract

There are three ways to get data from the internet into R:

* Make API queries “by hand” using httr.
* Scrape the web.
* Use an R package that wraps an API.

In this homework, I will practice with the first two methods.


#### Table of content
-   [Load data and packages](#load-data-and-packages)
-   [Make API queries by hand using httr](#make-api-queries-by-hand-using-httr)
-   [Scrape the web](#scrape-the-web)



Load data and packages
----------------------

[*Back to the top*](#abstract)

```{r,warning=FALSE, message=FALSE}
library(httr)
library(stringr)
library(tidyverse)
library(glue)
library(kableExtra)
library(knitr)
library(magrittr)
library(purrr)
library(xml2)
library(rvest)
```


Make API queries by hand using httr
-----------------------------------

[*Back to the top*](#abstract)

**(1)** I will write a function called `get_movie_TY` that get the information about movies from the Internet, this is [the website](http://www.imdb.com/) I got data from.
```{r}
get_movie_TY <- function(title,year){
  query_string <- glue("http://www.omdbapi.com/?t={title}&y={year}/?i=tt3896198&apikey=6f820a12")%>%
    str_replace_all(" ","-")
  movie_request <- GET(query_string)
  movie_content <- content(movie_request)
  return(movie_content)
}
```

**(2)** Use the function defined above, I will try to get some data about some of my favourite movies, and make a dataframe using `join` function.
```{r,warning=FALSE, message=FALSE}
frozen <- get_movie_TY("frozen","2013") 

flipped <- get_movie_TY("flipped","2010")

Despicable_Me <- get_movie_TY("Despicable Me","2010")

Big_Hero_6 <- get_movie_TY("Big Hero 6","2014")

Avatar <- get_movie_TY("Avatar","2009")

spirited_away <- get_movie_TY("spirited away","2001")

movie <- data.frame(Big_Hero_6)%>%
  full_join(data.frame(frozen))%>%
  full_join(data.frame(Despicable_Me))%>%
  full_join(data.frame(flipped))%>%
  full_join(data.frame(Avatar))%>%
  full_join(data.frame(spirited_away))
```

**(3)** Write this dataset to file, and try read it back in.
```{r, message=FALSE}
write_csv(movie, "movie.csv")
my_movie_data <-read_csv("movie.csv")
```

**(4)** Let's take a look at some columns of `my_movie_data`.
```{r}
movie_info <- my_movie_data[c("Title","Year", "Runtime", "Genre", "Director", "Actors","Language","Country")]
movie_info%>%kable()
```

**(5)** Let's compare the rating of these movies.
```{r}
my_movie_data[c("Title","Ratings.Value")]%>%kable()
```

Spirited Away has the highest rating score!

And here are the directors!
```{r}
my_movie_data[c("Title","Director")][6,2]%>%kable()
```


Scrape the web
--------------

[*Back to the top*](#abstract)

**(1)** Let's try to get some information about the movie *Big Hero 6* using `read_html`.
```{r}
my_url <- "http://www.imdb.com/title/tt2245084/?ref_=nv_sr_1"

my_movie <- read_html(my_url)
```

**(2)** Get the cast , the director and the rating of the movie.
```{r}
my_movie %>%
  html_nodes(css="#titleCast span.itemprop") %>%
  html_text()%>%
  collapse(sep = "\n")

my_movie %>%
  html_nodes(css=".credit_summary_item:nth-child(2)") %>%
  html_text()%>%
  str_replace_all("[\r\n]" , "")%>%
  str_replace_all("     " , " ")%>%
  str_trim()

my_movie %>%
  html_nodes(css=".ratingValue span") %>%
  html_text()%>%
  paste(collapse = '')

```

**(3)** I will now define a function `get_cast` to get the cast of movies, a function `get_director` to get the director of movies, and a function `get_rate` to get the rate of a movie. 
```{r}
get_cast <- function(movie_link){
  movie_link %>%
    read_html()%>%
    html_nodes(css="#titleCast span.itemprop")%>%
    html_text()%>%
    collapse(sep = "\n") %>%
    return()
}

get_director <- function(movie_link){
  movie_link %>%
    read_html()%>%
    html_nodes(css=".credit_summary_item:nth-child(2)")%>%
    html_text()%>%
    str_replace_all("[\r\n]" , "")%>%
    str_replace_all("     " , " ")%>%
    str_trim() %>%
    return()
}

get_rate <- function(movie_link){
  movie_link %>%
    read_html()%>%
    html_nodes(css=".ratingValue span")%>%
    html_text()%>%
    paste(collapse = '')%>%
    return()
}
```

**(4)** Try my function `get_cast` `get_director` and `get_rate` on the movie *Big Hero 6*.
```{r}
get_cast(my_url)
get_director(my_url)
get_rate(my_url)
```

**(5)** How about the movie *Frozen*?
```{r}
my_url_frozen <- "http://www.imdb.com/title/tt2294629/?ref_=fn_al_tt_1"
get_cast(my_url_frozen)
get_director(my_url_frozen)
get_rate(my_url_frozen)
```

*Transformers: The Last Knight*
```{r}
my_url_transformer <- "http://www.imdb.com/title/tt3371366/?ref_=nv_sr_1"
get_cast(my_url_transformer)
get_director(my_url_transformer)
get_rate(my_url_frozen)
```

[*Back to the top*](#abstract)
